---
title: "Linear Regression"
output: linearRegression

This File aims to explain what is Linear Regression under the hood. We are going the see the fundamentals of linear regrassion with a simple experiment. 

This is the data that we are going to use. This is the height, weight, age and sex of a set of individuous. 
---
```{r}
library(rethinking)
data(Howell1)
d <- Howell1

precis( d )
d2 <- d[ d$age >= 18 , ]
```

We want the predict the height, based on the others variables. This is the first step. We define the priors of this variable. 
We assume Normal distribution. 

```{r}

sample_mu <- rnorm( 1e4 , 178 , 20 ) 
sample_sigma <- runif( 1e4 , 0 , 50 )
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )
sample_mu <- rnorm( 1e4 , 178 , 100 ) 
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )

```

Calculando a posteriori a partir da prior 1. 



```{r}

mu.list <- seq( from=150, to=160 , length.out=100 )
sigma.list <- seq( from=7 , to=9 , length.out=100 )
post <- expand.grid( mu=mu.list , sigma=sigma.list )
post$LL <- sapply( 1:nrow(post) , function(i) sum(
dnorm( d2$height , post$mu[i] , post$sigma[i] , log=TRUE ) ) )
post$prod <- post$LL + dnorm( post$mu , 178 , 20 , TRUE ) +
dunif( post$sigma , 0 , 50 , TRUE )
post$prob <- exp( post$prod - max(post$prod) )

sample.rows <- sample( 1:nrow(post) , size=1e4 , replace=TRUE ,prob=post$prob )
sample.mu <- post$mu[ sample.rows ]
sample.sigma <- post$sigma[ sample.rows ]
plot( sample.mu , sample.sigma , cex=0.5 , col=col.alpha(rangi2,0.1) , xlab="mu" , ylab="sigma" , pch=16 )


```



```{r}
d3 <- d2
mu.list <- seq( from=150, to=170 , length.out=200 ) 
sigma.list <- seq( from=4 , to=20 , length.out=200 )
post2 <- expand.grid( mu=mu.list , sigma=sigma.list )
post2$LL <- sapply( 1:nrow(post2) , function(i) sum( dnorm( d3 , mean=post2$mu[i] , sd=post2$sigma[i] ,log=TRUE ) ) )
post2$prod <- post2$LL + dnorm( post2$mu , 178 , 20 , TRUE ) + dunif( post2$sigma , 0 , 50 , TRUE )
post2$prob <- exp( post2$prod - max(post2$prod) )

sample2.rows <- sample( 1:nrow(post2) , size=1e4 , replace=TRUE ,prob=post2$prob )
sample2.mu <- post2$mu[ sample2.rows ]
sample2.sigma <- post2$sigma[ sample2.rows ]
plot( sample2.mu , sample2.sigma , cex=0.5 , col=col.alpha(rangi2,0.1) , xlab="mu" , ylab="sigma" , pch=16 )
            
```


```{r}
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]
# define the average weight, x-bar
xbar <- mean(d2$weight)
# fit model
m4.3 <- quap(
alist(height ~ dnorm( mu , sigma ) ,
mu <- a + b*( weight - xbar ) ,
a ~ dnorm( 178 , 20 ) ,
b ~ dlnorm( 0 , 1 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data=d2 )
```

```{r}
plot( height ~ weight , data=d2 , col=rangi2 ) 
post <- extract.samples( m4.3 )
a_map <- mean(post$a)
b_map <- mean(post$b)
curve( a_map + b_map*(x - xbar) , add=TRUE )
```

Focus for the moment on a single weight value, say 50 kilograms. You can quickly make
a list of 10,000 values of µ for an individual who weighs 50 kilograms, by using your samples
from the posterior:

```{r}
post <- extract.samples( m4.3 )
mu_at_50 <- post$a + post$b * ( 50 - xbar )
```
 data), assuming the weight is 50 kg.
That’s good so far, but we need to repeat the above calculation for every weight value
on the horizontal axis, not just when it is 50 kg. We want to draw 89% intervals around the
average slope in Figure 4.6

Use link to generate distributions of posterior values for µ. The default behavior
of link is to use the original data, so you have to pass it a list of new horizontal axis
values you want to plot posterior predictions across.
What you’ve done so far is just use samples from the posterior to visualize the uncertainty
in µi,

```{r}

 mu <- link( m4.3 )
str(mu)
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
weight.seq <- seq( from=25 , to=70 , by=1 )
# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( m4.3 , data=data.frame(weight=weight.seq) )
str(mu)
# use type="n" to hide raw data
plot( height ~ weight , d2 , type="n" )
# loop over samples and plot each mu value
for ( i in 1:100 )
points( weight.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.1) )
```


